Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-12 paper.tex
some more wording

The project is a programming library, i.e. a collection of classes (in the
object-oriented sense) that users import into Python programs. Its use
therefore requires basic Python programming knowledge. No command-line
interface, let alone a GUI, is offered for non-programmer users. Instead,
interactive use is made possible by the Python interactive interpreter, and its
enhanced replacement IPython \citep{perez2007ipython}, which offers a
\textsc{matlab}-like working environment specifically designed for scientific
use.

@@

Within the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gaged from various indicators such as hundreds
of citations in scientific publications, successes in various machine learning challenges
(e.g., Kaggle), or significant statistics derived from our repositories
and mailing lists.  At the time of writing, the project is watched by 1309
people and forked 667 time on GitHub, and the mailing-list receives 300 mails
per month.
The version control logs for the development version of scikit-learn
% ddaa494c116e3c16bf032003c5cccbed851733d2
show 183 contributors to the project's source code and documentation
(after manual deduplication).

@@

The API is designed to adhere to the following broad principles:
\begin{description}
  \item[Inspectability.]
       Constructor parameters
       and parameter values determined by learning algorithms
       are stored as public attributes. All of them are documented in a consistent manner.
       % Gilles: Do we really need to bother the reader with such details?
       % (The Python language does not actually distinguish
       % between public and private attributes of objects,
       % but by convention, any identifier
       % prefixed with an underscore (\texttt{\_})
       % is considered ``a non-public part of the API''; \citealp{pythontut}.)
  \item[Non-proliferation of classes.]
       Learning algorithms are objects of custom classes,
       but datasets and hyperparameter names and values
       are represented as standard Python, NumPy or SciPy classes
       rather than new ones whenever feasible.
       % Gilles: I don't understand the phrase below
       % Custom classes adhere to Python conventions and protocols
       % for constructs such as iteration over the folds
       % in a $k$-fold cross validation.
       This should keep \sklearn easy to use
       and easy to combine with other libraries.
       % XXX do we need to elaborate more on this, or do we expect everyone
       % to be an experienced Java/adapter pattern hater?
  \item[Composition.]
       Many machine learning tasks are expressible
       as \textit{pipelines} of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parameterized on other algorithms.
       To be able to implement and use these algorithms in a simple way,
       all learners are usable with a consistent set of methods
       and a simple syntax is implemented
       for setting and retrieving parameters of sub-estimators.
       % XXX need to describe this!
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}

>>>>

The project is a programming library, i.e. a collection of classes -- in the
object-oriented sense -- that users import into Python programs. Using \sklearn
therefore requires basic Python programming knowledge. No command-line
interface, let alone a graphical user interface, is offered for non-programmer
users. Instead, interactive use is made possible by the Python interactive
interpreter, and its enhanced replacement IPython \citep{perez2007ipython},
which offer a \textsc{matlab}-like working environment specifically designed for
scientific use.

@@

Within the past few years, the project has proved valuable to an ever increasing
user base. Its popularity can be gaged from various indicators such as hundreds
of citations in scientific publications, successes in various machine learning
challenges (e.g., Kaggle), or significant statistics derived from our
repositories and mailing lists.  At the time of writing, the project is watched
by 1309 people and forked 667 times on GitHub, and the mailing-list receives
more than 300 mails per month. The version control logs for the development
version
% ddaa494c116e3c16bf032003c5cccbed851733d2
also count 183 contributors to the project (after manual deduplication).

@@

The API is designed to adhere to the following broad principles:
\begin{description}
  \item[Inspection.]
       Constructor parameters
       and parameter values determined by learning algorithms
       are stored as public attributes. All of them are documented in a consistent manner.
       % Gilles: Do we really need to bother the reader with such details?
       % (The Python language does not actually distinguish
       % between public and private attributes of objects,
       % but by convention, any identifier
       % prefixed with an underscore (\texttt{\_})
       % is considered ``a non-public part of the API''; \citealp{pythontut}.)
  \item[Non-proliferation of classes.]
       Learning algorithms are objects of custom classes,
       but datasets and hyperparameter names and values
       are represented as standard Python, NumPy or SciPy classes
       rather than new ones whenever feasible.
       % Gilles: I don't understand the phrase below
       % Custom classes adhere to Python conventions and protocols
       % for constructs such as iteration over the folds
       % in a $k$-fold cross validation.
       This should keep \sklearn easy to use
       and easy to combine with other libraries.
       % XXX do we need to elaborate more on this, or do we expect everyone
       % to be an experienced Java/adapter pattern hater?
  \item[Composition.]
       Many machine learning tasks are expressible
       as sequences or combinations of transformations to data.
       Some learning algorithms are also naturally viewed
       as meta-algorithms parameterized on other algorithms.
       To be able to implement and use these algorithms in a simple way,
       all learners are usable with a consistent set of methods
       and a simple syntax is implemented
       for setting and retrieving parameters of sub-estimators.
       % XXX need to describe this!
  \item[Sensible defaults.]
       Whenever an operation requires a user-defined parameter,
       a default and appropriate value is defined by the library.
       The default value should cause the operation to be performed
       in a sensible way (giving a baseline solution for the task at hand).
\end{description}
